Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-11-03 paper/intro.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@655 3fc257bb-8550-0410-81b3-a5c103da83f7

\item {\it Small, portable:} The entire MR-MPI library is a few
thousand lines of standard C++ code.  For parallel operation, the
program is linked with MPI, a standard message passing library
available on all distributed memory machines.  For serial operation, a
dummy MPI library (provided) can be substituted.  As a library, it can
be embedded in other codes~\cite{Titan} to enable them to perform
MapReduce operations.

\item {\it In-core or out-of-core operation:} Each MapReduce object
that a processor defines allocates ``pages'' of memory, where the page
size is determined by the user.  Typical MapReduce operations can be
performed using a few such pages (per processor).  If the data set
fits in a single page (per processor), the library performs its
operations in-core.  If the data set exceeds the page size, processors
each write to temporary disk files (on local disks or a parallel file
system) as needed and subsequently read from them.  This out-of-core
operation allows processing of data sets larger than the aggregate
memory of all the processors, i.e. up to the available aggregate disk
space.

\item {\it Flexible programmability:} An advantage of writing a
MapReduce program on top of MPI, is that the user program can invoke
MPI calls directly, if desired.  For example, one-line calls to
MPI\_Allreduce are often useful in determining the status of an
iterative graph algorithm, as described in Section~\ref{sec:graph}.
The library interface also provides a user data pointer as an argument
passed to all callback functions, so it is easy for the user program
to store ``state'' on each processor, accessible during the map and
reduce operations.  For example, various flags can be stored that
alter the operation of a map or reduce operation, as can richer data
structures that accumulate results.

\item {\it C++, C, and Python interfaces:} The C++ interface to the
MR-MPI library allows user programs to instantiate and then invoke
methods in one or more MapReduce objects.  The C interface allows the
library to be called from C or other high-level languages such as
Fortran.  The C interface also allows the library to be wrapped easily
by Python via the Python ``ctypes'' module.  The library can then be
called from a Python script, allowing the user to write {\it map()} and
{\it reduce()} callback functions in Python.  If a machine supports running
Python in parallel, a parallel MapReduce can also be run in this mode.

\item {\it No fault tolerance:} Current MPI implementations do not enable
easy detection of a dead processor or retrieval of the data it was
working on.  So like most MPI programs, a parallel program calling the
MR-MPI library will hang or crash if a processor goes away.  Unlike
Hadoop, and its HDFS file system which provides data redundancy,
the MR-MPI library reads and writes simple, flat files.  It can
use local per-processor disks, or a parallel file system, if
available, but these typically provide no data redundancy.

>>>>

\item {\it Small, portable:} The entire MR-MPI library is a few
thousand lines of standard C++ code.  For parallel operation, the
program is linked with MPI, a standard message passing library
available on all distributed-memory machines and many shared-memory
parallel machines.  For serial operation, a dummy MPI library
(provided) can be substituted.  As a library, MR-MPI can be embedded
in other codes~\cite{Titan} to enable them to perform MapReduce
operations.

\item {\it In-core or out-of-core operation:} Each MapReduce object
that a processor creates allocates ``pages'' of memory, where the page
size is determined by the user.  Typical MapReduce operations can be
performed using a few such pages (per processor).  If the data set
fits in a single page (per processor), the library performs its
operations in-core.  If the data set exceeds the page size, processors
each write to temporary disk files (on local disks or a parallel file
system) as needed and subsequently read from them.  This out-of-core
mode of operation allows processing of data sets larger than the
aggregate memory of all the processors, i.e. up to the aggregate
available disk space.

\item {\it Flexible programmability:} An advantage of writing a
MapReduce program on top of MPI, is that the user program can also
invoke MPI calls directly, if desired.  For example, one-line calls to
MPI\_Allreduce are often useful in determining the status of an
iterative graph algorithm, as described in Section~\ref{sec:graph}.
The library interface also provides a user data pointer as an argument
passed to all callback functions, so it is easy for the user program
to store ``state'' on each processor, accessible during the map and
reduce operations.  For example, various flags can be stored that
alter the operation of a map or reduce operation, as can richer data
structures that accumulate results.

\item {\it C++, C, and Python interfaces:} The C++ interface to the
MR-MPI library allows user programs to instantiate and then invoke
methods in one or more MapReduce objects.  The C interface allows the
same operations to be called from C or other high-level languages such
as Fortran.  The C interface also allows the library to be easily
wrapped by Python via the Python ``ctypes'' module.  The library can
then be called from a Python script, allowing the user to write {\it
map()} and {\it reduce()} callback functions in Python.  If a machine
supports running Python in parallel, a parallel MapReduce can also be
run in this mode.

\item {\it No fault tolerance or data redundancy:} Current MPI
implementations do not enable easy detection of a dead processor or
retrieval of the data it was working on.  So like most MPI programs, a
parallel program calling the MR-MPI library will hang or crash if a
processor goes away.  Unlike Hadoop, and its HDFS file system which
provides data redundancy, the MR-MPI library reads and writes simple,
flat files.  It can use local per-processor disks, or a parallel file
system, if available, but these typically provide no data redundancy.

### 2010-11-03 paper/mr.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@655 3fc257bb-8550-0410-81b3-a5c103da83f7

The basic datums stored and operated on by any MapReduce framweork are
key/value (KV) pairs.  In the MR-MPI library, individual keys or
values can be of any data type or length, or combinations of multiple
types (one integer, a string of characters, two integers and a double,
etc); they are simply treated as byte strings by the library.  A KV
pair always has a key; a KV's value may be NULL.  A related data type
is the key/multivalue (KMV) pair, where all values associated with the
same key are collected and stored contiguously as a multivalue, which
is just a longer byte string with an associated vector of lengths, one
integer length per value.  In this section, we assume the datums
operated on by each processor fit in local memory.

A typical MR-MPI program makes at least three calls to the MR-MPI
library, to perform {\it map()}, {\it collate()}, and {\it reduce()}
operations.  In a {\it map()}, zero or more key/value pairs are
generated by each processor.  Often this is done using data read from
files, but a {\it map()} may generate data itself or process existing
KV pairs to create new ones.  The KV pairs produced are stored locally
by each processor; a {\it map()} thus requires no inter-processor
communication.  Users call the library with a count of tasks to
perform and a pointer to a user function; the MR-MPI {\it map()}
operation invokes the user function multiple times as a callback.
Depending on which variant of {\it map()} is called, the user function
may be passed a file name, a chunk of bytes from a large file, a task
ID, or a KV pair.  Options for assigning map tasks to processors are
specified by the user and include assigning consecutive chunks of
tasks to processors, striding the tasks across processors, or using a
master-slave model that is useful when tasks have widely varying
workloads.

@@

The interface to various low-level operations is provided so
that a user's program can string them together in various ways to
produce interesting MapReduce algorithms.  For example, output from a
{\it reduce()} can serve as input to a subsequent {\it map()} or {\it
collate()}.  KV pairs from multiple MapReduce objects can be combined
to perform new sequences of {\it map()}, {\it collate()}, and
{\it reduce()} operations.

>>>>

The basic datums stored and operated on by any MapReduce framweork are
key/value (KV) pairs.  In the MR-MPI library, individual keys or
values can be of any data type or length, or combinations of multiple
types (one integer, a string of characters, two integers and a double,
etc); they are simply treated as byte strings by the library.  A KV
pair always has a key; its value may be NULL.  KV pairs are stored
within a MapReduce (MR) object.  A user program may create one or more
MR objects to implement an algorithm.  Various MapReduce operations
(map, reduce, etc) are invoked on an object and its KV pairs; KV pairs
can also be passed between and combined across objects.

A related data type in the MR-MPI library is the key/multivalue (KMV)
pair, where all values associated with the same key are collected and
stored contiguously as a multivalue, which is just a longer byte
string with an associated vector of lengths, one integer length per
value.  In this section, we assume the KV and KVM datums operated on
by each processor fit in local memory.

A typical MR-MPI program makes at least three calls to the MR-MPI
library, to perform {\it map()}, {\it collate()}, and {\it reduce()}
operations on a MR object it creates.  In a {\it map()}, zero or more
key/value pairs are generated by each processor.  Often this is done
using data read from files, but a {\it map()} may generate data itself
or process existing KV pairs to create new ones.  The KV pairs
produced are stored locally by each processor; a {\it map()} thus
requires no inter-processor communication.  Users call the library
with a count of tasks to perform and a pointer to a user function; the
MR-MPI {\it map()} operation invokes the user function multiple times
as a callback.  Depending on which variant of {\it map()} is called,
the user function may be passed a file name, a chunk of bytes from a
large file, a task ID, or a KV pair.  Options for assigning map tasks
to processors are specified by the user and include assigning
consecutive chunks of tasks to processors, striding the tasks across
processors, or using a master-slave model that is useful when tasks
have widely varying workloads.

@@

The interface to various low-level operations is provided so that a
user's program can string them together in various ways to construct
interesting MapReduce algorithms.  For example, output from a {\it
reduce()} can serve as input to a subsequent {\it map()} or {\it
collate()}.  Likewise, data can be partitioned across multiple
MapReduce (MR) objects, each with their own set of KV pairs.  KV pairs
from multiple MR objects can be combined to perform new sequences of
{\it map()}, {\it collate()}, and {\it reduce()} operations.

### 2010-11-03 paper/outcore.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@655 3fc257bb-8550-0410-81b3-a5c103da83f7

Since the MapReduce paradigm was designed to enable processing of
extremely large data sets, the MR-MPI library also allows for
``out-of-core'' processing, which is triggered when the KV or KMV
pairs owned by one or more processors do not fit in local memory.
When this occurs, a processor writes one or more temporary files to
disk, containing KV or KMV pairs, and reads them back in when
required.  Depending on the parallel machine's hardware configuration,
these files can reside on disks local to each processor, on the front
end (typically a NSF-mounted file system for a parallel machine), or
on a parallel file system/disk array.

@@

We now explain how the MapReduce operations described in the previous
section work in out-of-core mode.  The {\it map()} and {\it reduce()}
operations are relatively simple.  As a {\it map()} generates KV pairs
via the user function, a page of memory fills up, one KV pair at a
time.  When the page is full, it is written to disk.  If the source of
data for the {\it map()} operation is an existing set of KV pairs,
those datums are read, one page at a time, and a pointer to
each KV pair is given to the user function.  Similarly, a {\it
reduce()} reads one page of KMV pairs from disk, passes a pointer
to each pair, one at a time, to the user function that typically generates
new KV pairs.  The generated pairs fill up a new page, which is
written to disk when full, just as with the {\it map()} operation.
Thus for both a {\it map()} and {\it reduce()}, out-of-core disk files
are read and written sequentially, one page at a time, which requires
at most two pages of memory.

>>>>

Since the MapReduce paradigm was designed to enable processing of
extremely large data sets, the MR-MPI library also allows for
``out-of-core'' processing, which is triggered when the KV or KMV
pairs owned by one or more processors do not fit in local memory.
When this occurs, a processor writes one or more temporary files to
disk, containing KV or KMV pairs, and reads them back in when
required.  Depending on the parallel machine's hardware configuration,
these files can reside on disks local to each processor, on the front
end (typically a NFS-mounted file system for a parallel machine), or
on a parallel file system and its associated disk array.

@@

We now explain how the MapReduce operations described in the previous
section work in out-of-core mode.  The {\it map()} and {\it reduce()}
operations are relatively simple.  As a {\it map()} generates KV pairs
via the user function, a page of memory fills up, one KV pair at a
time.  When the page is full, it is written to disk.  If the source of
data for the {\it map()} operation is an existing set of KV pairs,
those datums are read, one page at a time, and a pointer to each KV
pair is given to the user function.  Similarly, a {\it reduce()} reads
one page of KMV pairs from disk, and passes a pointer to each pair,
one at a time, to the user function that typically generates new KV
pairs.  The generated pairs fill up a new page, which is written to
disk when full, just as with the {\it map()} operation.  Thus for both
a {\it map()} and {\it reduce()}, out-of-core disk files are read and
written sequentially, one page at a time, which requires at most two
pages of memory.

### 2010-11-03 paper/paper.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@655 3fc257bb-8550-0410-81b3-a5c103da83f7

We describe a parallel library written with message-passing (MPI)
calls that allows algorithms to be expressed in the MapReduce
paradigm.  This means the calling program does not need to include
explicit parallel code, but instead provides ``map'' and ``reduce''
functions that operate independently on elements of a data set
distributed across processors.  The library performs needed data
movement between processors.  We describe how typical MapReduce
funtionality can be implemented in an MPI context, and also in an
out-of-core manner for data sets that do not fit within the aggregate
memory of a parallel machine.  Our motivation for creating this
library was to enable graph algorithms to be written as MapReduce
operations, allowing processing of Terabyte-scale data sets.  We
outline MapReduce versions of several such algorithms: vertex ranking
via PageRank, triangle finding, connected component identification,
Luby's algorithm for maximally independent sets, and single-source
shortest-path calculation.  To test the algorithms on arbitrarily
large artificial graphs we generate randomized R-MAT matrices in
parallel; a MapReduce version of this operation is also described.
Performance and scalability results for the various algorithms are
presented for varying size graphs on a distributed-memory cluster.
For some cases, we compare the results with non-MapReduce algorithms,
different machines, and different MapReduce software, namely Hadoop.
Our open-source library is written in C++, is callable from C++, C,
Fortran, or scripting languages such as Python, and can run on any
parallel platform that supports MPI.

>>>>

We describe a parallel library written with message-passing (MPI)
calls that allows algorithms to be expressed in the MapReduce
paradigm.  This means the calling program does not need to include
explicit parallel code, but instead provides ``map'' and ``reduce''
functions that operate independently on elements of a data set
distributed across processors.  The library performs needed data
movement between processors.  We describe how typical MapReduce
funtionality can be implemented in an MPI context, and also in an
out-of-core manner for data sets that do not fit within the aggregate
memory of a parallel machine.  Our motivation for creating this
library was to enable graph algorithms to be written as MapReduce
operations, allowing processing of terabyte-scale data sets.  We
outline MapReduce versions of several such algorithms: vertex ranking
via PageRank, triangle finding, connected component identification,
Luby's algorithm for maximally independent sets, and single-source
shortest-path calculation.  To test the algorithms on arbitrarily
large artificial graphs we generate randomized R-MAT matrices in
parallel; a MapReduce version of this operation is also described.
Performance and scalability results for the various algorithms are
presented for varying size graphs on a distributed-memory cluster.
For some cases, we compare the results with non-MapReduce algorithms,
different machines, and different MapReduce software, namely Hadoop.
Our open-source library is written in C++, is callable from C++, C,
Fortran, or scripting languages such as Python, and can run on any
parallel platform that supports MPI.
